TIMES Ireland Model: Residential Sector

### 2021-05-13 manuscript.tex
13 May 2021 - Update finishing Intro 1st draft

\begin{figure}[!htbp]
 \centering
 \includegraphics[scale=0.4]{Figures/Non-ETS-Share.png} 
 \caption{Overview of Ireland's 2030 Climate Targets}
 \label{fig:IrelandClimatetargets}
\end{figure}

@@

\subsection{Behavioural and Financial barriers}
\label{Intro:barriers}
\par
In Ireland, the carbon tax was introduced in 2009 on petrol, diesel, oils and gas, and was extended to solid fuels (coal and peat) in 2013 and 2014. The rate per tonne of CO2 emissions was €15 from 2010 to 2011 and was
increased to €20 per tonne of CO2 emissions in 2012.
It applies exclusively on emissions in non-ETS sectors.
As shown in Figure 1, Carbon Tax receipts from 2010 to
2018 grew every year, reflecting strong economic growth,
and amounted to €431 million in 2018. Auto Diesel
accounts for 42% of carbon tax receipts, which is also
reflective of the current composition of the vehicle fleet
in Ireland with 60% of vehicles operating on diesel. It is
interesting to note that unlike carbon tax receipts from
The Irish government introduced Carbon tax in 2010, to influence fossil fuel demand, XX. The current Carbon tax is €33.50 and this will rise ti €100/tCO2 in 2030. The tax applies to all fossil fuel consumption across different economic sectors and while many European countries have taken this approach, as Sweden lead by example and succefully implanted a Carbon tax in 1993??. Current carbon tax? Swedens GNP?? Or PPP. The difficulty in carbon tax is that it can widen social inequality, whereby the lower income fossil fuelled homes who cannot afford upgrade and may already be in fuel poverty have to pay more on income on fuel and middle or high income homes can avoid the rising costs, by investing in alternative low carbon heating. The slow increase does afford time for lower income houses but with only expensive option currently available, some may have no option but to keep paying and receiving fuel subsidies. While the electricity network stretches to all corners of Ireland, The gas network is only reaches to 680,000 houses in towns and cities, with biomethance and renewable gas targets for 2030, X%. The district heating networks are new and currently only two in Dublin which reach about XX houses, 2030 target X%. large amount of data centres. Other future heating fuel options such as hydrogen, HVO, solar may also play a part. 
	ESOM can be used to start the conversation on..
Look at other ESOM, brief explain…
Other residential sectors..

\subsection{TIMES Ireland Model (TIM)}
\label{Intro:TIM}
\par
TIM was built to help Ireland provide a pathway to achieve 2030 and 2050 targets. The residential sector plays a key part of this and it was built from the SEAI 2018 energy balance, the BER database and CSO data. The data filtration process used in the residential sector, derived a lot of conversation of the BER database, a group know as Ireland Energy Wiki by the National Retrofitting Modelling Group grew from building the residential sector, questions about the filters, default values, assumptions around thermal equations (EC..) and uncertainties were used to improve the data before it was used in TIM . TIM has many internal project deadines, which required extensive meetings about the data, constraints and results as the modelling was progressing. The data and results were presented to 30 external reviewers to provide feedback, this process proved invaluable having different expertise with different presepectives enabled TIM to fill in gaps. TIM will also be open platform to help improve the model and provide transparency. TIM’s residential sector was built in parallel with LEAP’s residential sector, so that residential sector can easily simulate policy. The demand for results from TIM has increased significantly with increased government ambition, TIM will provide input into CAP 2021 and Ireland’s carbon budgets, working with DECC and the CCAC respectively will also provide feedback and improve the robustness of TIM. for these reasons TIM residential sector is a state of the art modelling platform.

\section{Literature Review}
\label{LitR}
``How should Ireland set GHG targets in the residential sector?” 
 % LIT REW INPUT DATA - EU \cite{Simoes2013TheProject}

>>>>

%Preliminary emissions data published by the EPA in January 2021 indicate that emissions from the residential sector increased by 9% in 2020, primarily due to the increase in home working and Covid-related restrictions on movement forcing people to spend more time at home. 
\begin{figure}[!htbp]
 \centering
 \includegraphics[scale=0.4]{Figures/Non-ETS-Share.png} 
 \caption{Overview of Ireland's 2030 Climate Targets}
 \label{fig:IrelandClimatetargets}
\end{figure}

@@

\subsection{Current Energy Policy}
\label{Intro:Policy}
The 2019 Climate Action Plan identified 29 actions that would reduce emissions from the sector by 40-45%.  These actions included:
Reducing fossil fuel use in the sector and moving to electricity to provide heating and hot water in buildings
500,000 building retrofits to a BER B2 / cost optimal equivalent or carbon equivalent
Installation of 600,000 heat pumps, with 400,000 of these in existing buildings
Development of district heating, including two initiatives of municipal scale with the potential to provide heat equivalent to the needs of about 50,000 homes
Complete the rollout of the Support Scheme for Renewable Heat
Increasing the number of Sustainable Energy Communities to 1,500

\subsubsection{Electricity}
\par
Decarbonising Electricity is key for Ireland, both the \textit{``stick"} (tax) approach which aims to deter investment in fossil fuel power generation, and the \textit{``carrot"} (subsidies) approach which aims to encourage renewable electricity is applied in Irish policy.
In Ireland, a €15/tCO2 carbon tax was introduced in 2009 to influence fossil fuel demand. The carbon tax receipts grew every year from 2010 to 2018 \cite{AnOffice}. Ireland's current carbon tax of €33.50/tCO2 will rise to €100/tCO2 in 2030. Many countries have taken this approach, especially in Europe, where Finland was the first country to introduce a carbon tax in 1990, while Sweden were second to introduce a carbon tax after Finland, the carbon tax rate is the highest in the world at €108.81/tCO2. A carbon price or tax now accounts for approx. 15\% of GHG emissions \cite{WorldBankGroup2020State2020}. 
The Renewable Energy Feed-in Tariff (REFIT) schemes were designed to incentives investment in renewable electricity to achieve legally binding renewable targets under 2009/28/EC. REFIT operates  by  guaranteeing  new renewable generation and biomass a minimum price for electricity over a 15 year period. The last REFIT schemes closed in December 2015. In 2020 REFIT was replaced by the Renewable Electricity Support Scheme (RESS), which supports to renewable electricity projects to help Ireland achieve 2030 climate targets, especially the 70\% renewable electricity target. RESS delivers for communities to participate in renewable energy projects and increasing technology diversity by broadening the renewable electricity technology mix, where almost 800 MW of solar has been approved in RESS-1 \cite{Eirgrid2020RenewableResults}, this builds alot on the current solar capacity of 16 MW, while making significant progress towards the CAP2019 target of 1500 MW by 2030.

\subsubsection{Heat}
Domestic supports for renewable heat or energy efficiency is through SEAI grants,\begin{itemize}
    \item {Insulation}
    \subitem {Attic}
    \subitem {Cavity Wall}
    \subitem {Internal - Dry Lining}
    \subitem {External}
    \item {Heat Pumps}
    \subitem {Air to Water}
    \subitem {Ground Source to Water}
    \subitem {Exhaust Air to Water}
    \subitem {Water to Water}
    \subitem {Air to Air}
    \item {Heating Controls}
    \item {Solar Water Heating}
\end{itemize}

This report is a study of potential design options for a Renewable Heat Incentive (RHI) f

-  critical for the development of renewable gas projects in Ireland.European Commission’s decision on State aid for the Support Scheme for Renewable Heat (SA.50807).
GNI Vision 2050 - 20\% renewable gas by 2030
While the electricity network reaches all residential dwellings, The gas network reaches 40\% or 680,000 dwellings in towns and cities. The 2030 gas decarbonisation targets, set by GNI  
and  with biomethance and renewable gas targets for 2030, X\%. The district heating networks are new and currently only two in Dublin which reach about XX houses, 2030 target X%. large amount of data centres. Other future heating fuel options such as hydrogen, HVO, solar may also play a part. 
Gas Networks Ireland (GNI) has a strategic plan to achieve 20% Renewable Gas on the network by 2030 
Further investigate decarbonisation options such as hydrogen vehicles, biomethane and AD
substitutes for natural gas - See \ref{fig:CAP2019:MACC} shows biomethane as the most expensive options..

\subsection{TIMES Ireland Model (TIM)}
\label{Intro:TIM}
\par
TIM was built to help Ireland provide a pathway to achieve 2030 and 2050 targets. The residential sector plays a key part of this and it was built from the SEAI 2018 energy balance, the BER database and CSO data. The data filtration process used in the residential sector, derived a lot of conversation of the BER database, a group know as Ireland Energy Wiki by the National Retrofitting Modelling Group grew from building the residential sector, questions about the filters, default values, assumptions around thermal equations (EC..) and uncertainties were used to improve the data before it was used in TIM . TIM has many internal project deadines, which required extensive meetings about the data, constraints and results as the modelling was progressing. The data and results were presented to 30 external reviewers to provide feedback, this process proved invaluable having different expertise with different presepectives enabled TIM to fill in gaps. TIM will also be open platform to help improve the model and provide transparency. TIM’s residential sector was built in parallel with LEAP’s residential sector, so that residential sector can easily simulate policy. The demand for results from TIM has increased significantly with increased government ambition, TIM will provide input into CAP 2021 and Ireland’s carbon budgets, working with DECC and the CCAC respectively will also provide feedback and improve the robustness of TIM. for these reasons TIM residential sector is a state of the art modelling platform. 
%GRAPH OF IRELAND DECARBONISATION TARGETS  VS OTHER COUNTRIES - DIFFERENT INTRASTATE and RESOURCE AT BIDENS ENERGY SUBMIT - DENMAKR TOP 2018, 
	ESOM can be used to start the conversation on..
Look at other ESOM, brief explain…
Other residential sectors..
CAP2019 - Choosing the Pathways which Create the Least Burden
and Offer the Most Opportunity for Ireland
\section{Literature Review}
\label{LitR}
``How should Ireland set GHG targets in the residential sector?” 
 % LIT REW INPUT DATA - EU \cite{Simoes2013TheProject}
